\documentclass[addpoints,12pt]{exam}
\printanswers
\usepackage{amsmath,amssymb,complexity,amsthm}
\usepackage{graphicx}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\SD}{{\sf SD}}
\newcommand{\N}{{\mathbb{N}}}

\begin{document}
\hrule
\vspace{3mm}
\noindent
{\sf IITM-CS6014 : Advanced Theory of Computation  \hfill Author: Karthik Abinav}
\vspace{3mm}\\
\noindent
{\sf Problem Set \#4}
\vspace{3mm}
\hrule
{\small
\begin{itemize}
\item Rules of the game remain the same. Submissions must be single file in \LaTeX~format at the upload link set up in cse moodle page of the course. 
\end{itemize}}
\hrule

\begin{questions}
\question[15]
Argue whether the following functions qualify to be called as a
\emph{resource} according to Blum's resource axioms.
\begin{parts}
\part[3]$ {\sf VALUE}(M,x) = \begin{cases} 
			 0 & \text{ if $M$ rejects $x$}, \\
			 1 & \text{ if $M$ accepts $x$}, \\
			 \text{undefined} & \text{ if $M$ does not halts on
			 $x$}
		 	\end{cases}
$
\newline 
\textbf{Answer} \newline
The above resource function satisfies Blums first axiom i.e. it is defined only when $M$ halts on $x$. However, given a $M$,$x$,$k$ , it is not decidable
to check if VALUE(M,x) = k. Hence it is not a resource.

Consider $R= \{(M,x):VALUE(M,x)=k\}$

Suppose $R$ was decidable via a total TM $N$, then $MP$ can be decided by giving $(M,x,1)$ as input to $N$. But, $MP$ is not decidable and hence $R$ is not decidable.


\part[5] A \emph{turn} of tape head is defined as movement of tape head from
$L \to R$ or $R \to L$. 

$ {\sf TURN}(M,x) = \begin{cases}
			 \text{No. of turns that tape head makes when $M$ runs
			 on $x$} & \text{ if $M$ halts on $x$} \\
			 \text{undefined} & \text{ otherwise}
			\end{cases}
$
\newline 
\textbf{Answer} \newline
The above resource function satisfies Blums first axiom i.e. it is defined only when $M$ halts on $x$. However, given a $M$,$x$,$k$ , it is not decidable
to check if VALUE(M,x) = k. Hence it is not a resource.

Consider $R= \{(M,x):VALUE(M,x)=k\}$

Suppose $R$ was decidable via a total TM $N$, then $MP$ can be decided by giving $(M,x,1)$ as input to $N$. But, $MP$ is not decidable and hence $R$ is not decidable.



\part[5] $ {\sf COUNT}(M,x) = \begin{cases}
			 \text{No. of times $M$ vists a state $q$}
			 & \text{ if $M$ halts on $x$} \\
			 \text{undefined} & \text{ otherwise}
			\end{cases}
$
\end{parts}

\question[5] 
Show that if $\NTIME(n) = \DTIME(n)$ then $\P = \NP$. (Padding !!)

\question[10]
 Space Hierarchy theorem implies the following: For any $k >
0$, There is a language in $\DSPACE(n^{k+1})$ that is not in
$\DSPACE(n^k)$. Use this and a padding argument to show that: $\P
\neq \DSPACE(n)$. (6pts) \\ (Note that we do not know the containment
in either direction.) \\ You can do this in two steps:
\begin{parts}
\part For every language $L$ define, $L_{pad} = \{x01^{|x|^2}~:~x
      \in L\}$.\\ Argue that $L_{pad}$ is in $\P$ $\implies$ $L \in
      \P$.
\part Show an $L_{pad}$ which is in $\DSPACE(n)$ but whose
      corresponding $L$ is not in $\DSPACE(n)$.
\end{parts}

\question[5]
Show that if $\SAT \in \NP \cap \co\NP$ then $\NP = \co\NP$. (Definitions !)

\question[5]
If $L, L'$ are in $\NP$, then show that $L \cup L'$, $L \cap L'$ are in $\NP$. (Definitions !)
  
\question[5]
If $L, L'$ are in $\NP \cap \co\NP$, then show that $L \oplus  L'$ defined as 
\[ L \oplus L' = \{ x : x \textrm{ is in one of $L$ or $L'$ but not both.} \} \]
is in $\NP \cap \co\NP$. (Definitions !)

\question[15]
Consider the following language: PRIMES $= \{ n ~|~ n
\textrm{ is a prime } \}$ where the input $n$ is in binary.  Without
using the known result that PRIMES is in $\P$, solve the following:
\begin{parts}
\part[5] Show that PRIMES is in $\co\NP$.
\part[10] Here is Lucas test for primality (you dont need prove it) : $n$ is prime if and only if there is an integer $a \in \{2, \ldots,
  n-1\}$ with $a^{n-1} \equiv 1 \mod n$, and for every prime factor $q$ of $n-1$ : $a^{\frac{n-1}{q}} \not\equiv 1 \mod n$. Use this test to show that PRIMES is in $\NP$.
\end{parts}
Hence conclude that PRIMES is in $\NP$ $\cap$ $\co\NP$.

%\question
%Show that $\DSPACE(n^5) \subseteq \NSPACE(n^2)$.

\question[10] Prove that reachability in undirected forests (a possibly disconnected acyclic undirected graph) can be solved in $\log$-space. That is, given $(T,s,t)$ where $T$ is an undirected forest, it can be tested in $\log$-space whether $s$ is connected to $t$ by a path.

\question[18]
Let $\E = \bigcup_{c > 0} \DTIME(2^{cn})$ and $\NE = \bigcup_{c > 0} \NTIME(2^{cn})$. A set $A$ is called {\em sparse} if there is a polynomial $p$,
    such that $|\{ x \in A : |x| = n \}| \le p(n)$. A set $A$ is called
    {\em tally set} if $A \subseteq \{1\}^*$.  Prove that following
    are equivalent.
  \begin{enumerate}
  \item Restricted to tally sets $\NP = \P$. That is all tally sets in $\NP$ are in $\P$.
  \item Restricted to sparse sets $\NP = \P$. That is all sparse sets
    in $\NP$ are in $\P$.
  \item $\E = \NE$
  \end{enumerate}
  Hence conclude that $\E \ne \NE \implies \P \ne \NP$.
  
  \emph{Hint : Try for (b) $\implies$ (a) $\implies$ (c) $\implies$ (b). 
  For the second implication : consider the language $L_{tally} = \{1^{2^{c|x|}} : x \in L\}$.
  This will not work, but a slight modification of this language which includes some more information about 
  $x$ will work !. \\ For the third implication, consider the language 
\[ L_{order} = \{(k,i,c): \textrm{the $i^{th}$ bit of the $k^{th}$ string(in lex order) in L is $c$} \} \] }

\question[7] 
Imagine a world in which $\P = \NP$. Now show that there is a
polynomial time algorithm which given a Boolean formula $\phi$ produces a satisfying assignment for $\phi$ if $\phi$ is satisfiable.(Hint : Use queries to SAT).

\end{questions}



\end{document}
